---
title: "Activity 4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 1 - Activity 4

## Task 2: Load the necessary packages

```{r loading libraries,message=FALSE,warning=FALSE}
library(tidyverse)
library(tidymodels)
library(GGally)
library(ggfortify)
```


## Task 3: Load the data

```{r Loading Data}
evals<-read_tsv("https://www.openintro.org/data/tab-delimited/evals.txt",show_col_types = FALSE)
```

1) This is an experimental study for me because the data was collected from students to test the hypothesis whether beauty leads directly to affect the score evaluations. We can answer the question if it phrased as whether or not beauty affects the score evaluations.


```{r Distribution of Score}
plot1<-evals%>%ggplot(aes(x=score))+
  geom_histogram(fill="red",color="black",bins = 30)+
  labs(title = "Distribution of Score")+
  theme(plot.title = element_text(hjust = 0.5))
plot1
```

2) The distribution of the variable "Score" appears to be Left-skewed. It indicates that the majority of students gave the professors a more than average rating. It is not what I expected; I expected it to be slightly right-skewed.


3) I chose the variables gender and bty avg to show their relationship. I chose a box plot to visualize their relationship, and the below visualization shows that female professors have a slightly higher average beauty score than male professors.

```{r Plot2}
plot2<-evals%>%ggplot(aes(x=gender,y=bty_avg,fill=gender))+
  geom_boxplot()+
  labs(title = "Relationship between Gender and Beauty Average",
       x="Gender of the Professor",
       y="Beauty average score")+
  theme(plot.title = element_text(hjust = 0.5))

plot2
```

## Task 4: Pairwise relationships

```{r ggpairs}
evals_bty<-select(evals,c("bty_f1lower","bty_f1upper","bty_f2upper","bty_m1lower","bty_m1upper","bty_m2upper","bty_avg"))

evals_ggpairs<-evals_bty %>% ggpairs()

evals_ggpairs
```

4)  
  * bty_avg has a higher correlation with rest of the beauty variables and positively linear relation with bty_m1upper

  * bty_f1lower has the highest correlation with bty_avg and bty_m1upper
    
  * bty_f1upper has the highest correlation with bty_avg and bty_m1upper
    
  * bty_f2upper has positive linear relationship with bty_avg
    
  * bty_m1lower has positive linear relationship with bty_avg
    
  * bty_m1upper has positive linear relationship with bty_avg
    
  * bty_m2upper has positive linear relationship with bty_avg
    
    
5) I don't think it makes sense to include all the variables in the model, because it would be a complex model with many variables.

6) I think including bty_avg variable in the model will be a good decision since it has a positive linear relationship with all other beauty variables.


## Task 5: Multiple linear regression: one quantitative predictor, one qualitative predictor

```{r}
m_bty_gen <- lm(score ~ bty_avg+gender, data = evals)
tidy(m_bty_gen)
```


```{r}
# Creating the diagnostic plots using the function plot() from ggfortify package 
par(mfrow = c(2, 2))
plot(m_bty_gen)
```


    
7)  From the above diagnostic plots we can see that

    * The linear relationship assumption is met very well as the residual vs fitted plot does not follow        any particular patter and spread randomly
    * The normality assumption has met satisfactorily as the residuals follow along the line in q-q plot
    * The constant variance assumption also met reasonably as the residuals spread is equal along the           horizantal line
    
8) The bty_avg looks like a significant predictor of the score. The addition of gender has changed the parameter estimate of bty_avg from 3.88 to 3.75.


9) 
$$
  \begin{aligned}
\widehat{\texttt{score}} &= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times (1) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg}+ \hat{\beta}_2
\end{aligned}
$$