---
title: "Activity 4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 1 - Activity 4

## Task 2: Load the necessary packages

```{r loading libraries,message=FALSE,warning=FALSE}
library(tidyverse)
library(tidymodels)
library(GGally)
```


## Task 3: Load the data

```{r Loading Data}
evals<-read_tsv("https://www.openintro.org/data/tab-delimited/evals.txt",show_col_types = FALSE)
```

1) This is an experimental study for me because the data was collected from students to test the hypothesis whether beauty leads directly to affect the score evaluations. We can answer the question if it phrased as whether or not beauty affects the score evaluations.


```{r Distribution of Score}
plot1<-evals%>%ggplot(aes(x=score))+
  geom_histogram(fill="red",color="black",bins = 30)+
  labs(title = "Distribution of Score")+
  theme(plot.title = element_text(hjust = 0.5))
plot1
```

2) The distribution of the variable "Score" appears to be Left-skewed. It indicates that the majority of students gave the professors a more than average rating. It is not what I expected; I expected it to be slightly right-skewed.


```{r Plot2}
plot2<-evals%>%ggplot(aes(x=gender,y=bty_avg,fill=gender))+
  geom_boxplot()+
  labs(title = "Relationship between Gender and Beauty Average",
       x="Gender of the Professor",
       y="Beauty average score")+
  theme(plot.title = element_text(hjust = 0.5))

plot2
```

