---
title: "Activity 4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 1 - Activity 4

## Task 2: Load the necessary packages

```{r loading libraries,message=FALSE,warning=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidymodels))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(ggfortify))
```


## Task 3: Load the data

```{r Loading Data}
evals<-read_tsv("https://www.openintro.org/data/tab-delimited/evals.txt",show_col_types = FALSE)
```

1) This is an experimental study for me because the data was collected from students to test the hypothesis whether beauty leads directly to affect the score evaluations. We can answer the question if it phrased as whether or not beauty affects the score evaluations.


```{r Distribution of Score}
plot1<-evals%>%ggplot(aes(x=score))+
  geom_histogram(fill="red",color="black",bins = 30)+
  labs(title = "Distribution of Score")+
  theme(plot.title = element_text(hjust = 0.5))
plot1
```

2) The distribution of the variable "Score" appears to be Left-skewed. It indicates that the majority of students gave the professors a more than average rating. It is not what I expected; I expected it to be slightly right-skewed.


3) I chose the variables gender and bty avg to show their relationship. I chose a box plot to visualize their relationship, and the below visualization shows that female professors have a slightly higher average beauty score than male professors.

```{r Plot2}
plot2<-evals%>%ggplot(aes(x=gender,y=bty_avg,fill=gender))+
  geom_boxplot()+
  labs(title = "Relationship between Gender and Beauty Average",
       x="Gender of the Professor",
       y="Beauty average score")+
  theme(plot.title = element_text(hjust = 0.5))

plot2
```

## Task 4: Pairwise relationships

```{r ggpairs}
evals_bty<-select(evals,c("bty_f1lower","bty_f1upper","bty_f2upper","bty_m1lower","bty_m1upper","bty_m2upper","bty_avg"))

evals_ggpairs<-evals_bty %>% ggpairs()

evals_ggpairs
```

4)  
  * bty_avg has a higher correlation with rest of the beauty variables and positively linear relation with bty_m1upper

  * bty_f1lower has the highest correlation with bty_avg and bty_m1upper
    
  * bty_f1upper has the highest correlation with bty_avg and bty_m1upper
    
  * bty_f2upper has positive linear relationship with bty_avg
    
  * bty_m1lower has positive linear relationship with bty_avg
    
  * bty_m1upper has positive linear relationship with bty_avg
    
  * bty_m2upper has positive linear relationship with bty_avg
    
    
5) I don't think it makes sense to include all the variables in the model, because it would be a complex model with many variables.

6) I think including bty_avg variable in the model will be a good decision since it has a positive linear relationship with all other beauty variables.


## Task 5: Multiple linear regression: one quantitative predictor, one qualitative predictor

```{r}
m_bty_gen <- lm(score ~ bty_avg+gender, data = evals)
tidy(m_bty_gen)
```


```{r}
# Creating the diagnostic plots using the function plot() from ggfortify package 
par(mfrow = c(2, 2))
plot(m_bty_gen)
```


    
7)  From the above diagnostic plots we can see that

    * The linear relationship assumption is met very well as the residual vs fitted plot does not follow        any particular patter and spread randomly
    * The normality assumption has met satisfactorily as the residuals follow along the line in q-q plot
    * The constant variance assumption also met reasonably as the residuals spread is equal along the           horizantal line
    
8) The bty_avg looks like a significant predictor of the score. The addition of gender has changed the parameter estimate of bty_avg from 3.88 to 3.75.


9) 

$$
  \begin{aligned}
\widehat{\texttt{score}} &= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times (1) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg}+ \hat{\beta}_2
\end{aligned}
$$


10) For two genders received the same beauty avg score, The male professors are likely to receive a higher course evaluation score.


```{r}
# Model to predict the score with help of bty_avg and rank

m_bty_rank<-lm(score ~ bty_avg+rank, data = evals)
tidy(m_bty_rank)
```

11) R appears to handle categorical variables with more than 2 levels by removing one level from the total number of levels. Here since there are 3 levels, R is using only 2 of them by removing teaching and recodes the levels by creating a contrast matrix.

* If Rank is tenure track then the ranktenure track is coded as 1 and ranktenured as 0. 
* If Rank is tenured then the ranktenure track is coded as 0 and ranktenured as 1.
* If Rank is teaching then both the ranktenure track and ranktenured are coded as 0.


# Activity 4 - Day 2

## Task 2: bty_avg * gender interaction

```{r}
# model predicting score with help of bty_avg, gender and their interaction

m_int <- lm(score ~ bty_avg * gender, data = evals)
tidy(m_int)
```


1) Since R recodes the categorical variable into a indicator variable of binary values. Here, the gender variale is recoded as gendermale which takes value '0' for female and '1' for male professors. The interaction of bty_avg and gender is bty_avg:gendermale.

2a) For Male professors

$$
\begin{aligned}
\widehat{\texttt{score}} &= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times \texttt{gendermale}+ \hat{\beta}_3 \times \texttt{bty\\_avg:gendermale} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times (1)+ \hat{\beta}_3 \times \texttt{bty\\_avg:gendermale} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg}+ \hat{\beta}_2 + \hat{\beta}_3 \times \texttt{bty\\_avg:gendermale}
\end{aligned}
$$


2b) For Female Professors


$$
\begin{aligned}
\widehat{\texttt{score}} &= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times \texttt{gendermale}+ \hat{\beta}_3 \times \texttt{bty\\_avg:gendermale} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times (0)+ \hat{\beta}_3 \times \texttt{bty\\_avg:gendermale} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg}+ \hat{\beta}_3 \times \texttt{bty\\_avg:gendermale}
\end{aligned}
$$


3) When two professors who received the same beauty rating, the male professors tends to have the higher course evaluation score


```{r}
# checking the model fitof m_int and m_bty_gen

glance(m_int)

glance(m_bty_gen)
```

4) The interaction model is performing better than the model with jus the variables. I used the R-squared value and p-value to compare the model fit of both the models because the R-squared value tells us variance explained byeach model to predict the evaluation score.


## Task 3: bty_avg * rank interaction

```{r}
m_int_rank <- lm(score ~ bty_avg * rank, data = evals)
tidy(m_int_rank)
```


In this model since there are more than two levels in the categorical variable rank, r has recoded it to two variables by creating a contrast matrix and the same in the interaction of those two with the bty_avg variable. So, we have two interaction variables. Below here is how the model looks based on their ranks respectively

a) How the model looks irrespective of Rank



$$
\widehat{\texttt{score}} = \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times \texttt{ranktenure track}+ \hat{\beta}_3 \times \texttt{ranktenured} + \hat{\beta}_4 \times \texttt{bty_avg:ranktenure track}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured}
$$


b) If Rank of the professor is tenure track


$$
\begin{aligned}
\widehat{\texttt{score}} &= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times \texttt{ranktenuretrack}+ \hat{\beta}_3 \times \texttt{ranktenured}+ \hat{\beta}_4 \times \texttt{bty_avg:ranktenuretrack}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times (1)+ \hat{\beta}_3 \times (0)+ \hat{\beta}_4 \times \texttt{bty_avg:ranktenuretrack}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 + \hat{\beta}_4 \times \texttt{bty_avg:ranktenuretrack}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured}
\end{aligned}
$$

c) If Rank of the professor is tenured


$$
\begin{aligned}
\widehat{\texttt{score}} &= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times \texttt{ranktenure track}+ \hat{\beta}_3 \times \texttt{ranktenured}+ \hat{\beta}_4 \times \texttt{bty_avg:ranktenure track}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times (0)+ \hat{\beta}_3 \times (1)+ \hat{\beta}_4 \times \texttt{bty_avg:ranktenure track}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_3 + \hat{\beta}_4 \times \texttt{bty_avg:ranktenure track}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured}
\end{aligned}
$$

d) If Rank of the professor is teaching


$$
\begin{aligned}
\widehat{\texttt{score}} &= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times \texttt{ranktenure track}+ \hat{\beta}_3 \times \texttt{ranktenured}+ \hat{\beta}_4 \times \texttt{bty_avg:ranktenure track}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_2 \times (0)+ \hat{\beta}_3 \times (0)+ \hat{\beta}_4 \times \texttt{bty_avg:ranktenure track}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured} \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times \texttt{bty\\_avg} + \hat{\beta}_4 \times \texttt{bty_avg:ranktenure track}+ \hat{\beta}_5 \times \texttt{bty_avg:ranktenured}
\end{aligned}
$$


* For two professors who received the same beauty_avg rating, then the score would be higher in the order of ranking Teaching > tenure track > tenured


```{r}
# Checking the model fit of the models with and without the interaction of bty_avg and rank

glance(m_int_rank)
glance(m_bty_rank)
```

* Looks like the The model with the interaction of bty_avg and rank is performing better than the model without interaction based on the r-squared value and p-value